Notes: New gaugings in 3d from three-form

### 2022-07-22 Notes_3dH3.tex
New section 4 with notes on S3xS3xS1.

% ---- MATHS ----
\usepackage{amsmath,amssymb,amsfonts,dsfont}
\usepackage{mathrsfs}
\usepackage{physics}
\usepackage{ytableau}
\ytableausetup{boxsize=1.1em,centertableaux}
\usepackage{stmaryrd}
\usepackage{nicefrac}
\allowdisplaybreaks[1]
% \usepackage{bbold}
\usepackage{cases}
\usepackage{bm}

@@

% ---- COMMENTS ----
\newcommand{\ce}[1]{\marginpar{\parbox{\marginparwidth}{\boldmath $\Longleftarrow$}}
{\boldmath\bfseries #1}}

@@

\paragraph{\boldmath $\SO(8)_{\rm IIA}$ singlets}
Under $\E_{7(7)}\rightarrow\SL(8)_{\rm IIA}$:
\begin{equation}
  \begin{aligned}
    \bm{133} & \longrightarrow \bm{63} \oplus \bm{70},\\
    \bm{1539} & \longrightarrow \bm{63} \oplus \bm{378} \oplus \bm{\bar{378}} \oplus \bm{720}.
  \end{aligned}
\end{equation}
No singlet other that $\bm{1}$.

\paragraph{\boldmath $\SO(7)$ singlets}
Under $\E_{7(7)}\rightarrow\SL(7)\times\R_{+}$:
\begin{equation}
  \begin{aligned}
    \bm{133} & \longrightarrow \bm{1}_{0} \oplus \ldots,\\
    \bm{1539} & \longrightarrow \bm{1}_{0} \oplus \ldots
  \end{aligned}
\end{equation}
where only representations featuring $\SO(7)$ singlets have been specified.

\paragraph{\boldmath $\SO(6)\times\R_{+}$ singlets}
Under $\E_{7(7)}\rightarrow\SL(6)\times\SL(2)\times\R_{+}$:
\begin{equation}
  \begin{aligned}
    \bm{133} & \longrightarrow (\bm{1},\bm{1})_{0} \oplus \ldots,\\
    \bm{1539} & \longrightarrow 2\times(\bm{1},\bm{1})_{0} \oplus \ldots\\
  \end{aligned}
\end{equation}
where only representations featuring $\SO(6)\times\R_{+}$ singlets have been specified.

\paragraph{\boldmath $\SO(5)\times\SO(3)$ singlets}
Under $\E_{7(7)}\rightarrow\SL(5)\times\SL(3)\times\R_{+}$:
\begin{equation}
  \begin{aligned}
    \bm{133} & \longrightarrow (\bm{1},\bm{1})_{0} \oplus \ldots,\\
    \bm{1539} & \longrightarrow 2\times(\bm{1},\bm{1})_{0} \oplus \ldots\\
  \end{aligned}
\end{equation}
where only representations featuring $\SO(5)\times\SO(3)$ singlets have been specified.

\paragraph{\boldmath $\SO(4)\times\SO(4)$ singlets}
Under $\E_{7(7)}\rightarrow\SL(4)\times\SL(4)\times\R_{+}$:
\begin{equation}
  \begin{aligned}
    \bm{133} & \longrightarrow (\bm{1},\bm{1})_{0} \oplus (\bm{1},\bm{1})_{4} \oplus (\bm{1},\bm{1})_{-4} \oplus \ldots,\\
    \bm{1539} & \longrightarrow 2\times(\bm{1},\bm{1})_{0} \oplus \ldots\\
  \end{aligned}
\end{equation}
where only representations featuring $\SO(4)\times\SO(4)$ singlets have been specified.

\paragraph*{} The singlets always come from the $\bm{63}$ and $\bm{720}$ of $\SL(8)_{\rm IIB}$ (and $\bm{70}$ for $\SL(4)\times\SL(4)\times\R_{+}$).

>>>>

% ---- MATHS ----
\usepackage{amsmath,amssymb,amsfonts,dsfont}
\usepackage{mathrsfs}
\usepackage{physics}
\usepackage{ytableau}
\ytableausetup{boxsize=1.1em,centertableaux}
\usepackage{stmaryrd}
\usepackage{nicefrac}
\allowdisplaybreaks[1]
% \usepackage{bbold}
\usepackage{cases}
\usepackage{bm}
\usepackage{extarrows}

@@

% ---- COMMENTS ----
\newcommand{\ce}[1]{\marginpar{\parbox{\marginparwidth}{\boldmath $\Longleftarrow$}}
{\boldmath\bfseries (ce: #1)}}

@@

% \paragraph{\boldmath $\SO(8)_{\rm IIA}$ singlets}
Under $\E_{7(7)}\rightarrow\SL(8)_{\rm IIA}$:
\begin{equation}
  \begin{aligned}
    \bm{133} & \longrightarrow \bm{63} \oplus \bm{70},\\
    \bm{1539} & \longrightarrow \bm{63} \oplus \bm{378} \oplus \bm{\bar{378}} \oplus \bm{720}.
  \end{aligned}
\end{equation}
No $\SO(8)_{\rm IIA}$ singlet other that $\bm{1}$.

% \paragraph{\boldmath $\SO(7)$ singlets}
% Under $\E_{7(7)}\rightarrow\SL(7)\times\R_{+}$:
% \begin{equation}
%   \begin{aligned}
%     \bm{133} & \longrightarrow \bm{1}_{0} \oplus \ldots,\\
%     \bm{1539} & \longrightarrow \bm{1}_{0} \oplus \ldots
%   \end{aligned}
% \end{equation}
% where only representations featuring $\SO(7)$ singlets have been specified.

% \paragraph{\boldmath $\SO(6)\times\R_{+}$ singlets}
% Under $\E_{7(7)}\rightarrow\SL(6)\times\SL(2)\times\R_{+}$:
% \begin{equation}
%   \begin{aligned}
%     \bm{133} & \longrightarrow (\bm{1},\bm{1})_{0} \oplus \ldots,\\
%     \bm{1539} & \longrightarrow 2\times(\bm{1},\bm{1})_{0} \oplus \ldots\\
%   \end{aligned}
% \end{equation}
% where only representations featuring $\SO(6)\times\R_{+}$ singlets have been specified.

% \paragraph{\boldmath $\SO(5)\times\SO(3)$ singlets}
% Under $\E_{7(7)}\rightarrow\SL(5)\times\SL(3)\times\R_{+}$:
% \begin{equation}
%   \begin{aligned}
%     \bm{133} & \longrightarrow (\bm{1},\bm{1})_{0} \oplus \ldots,\\
%     \bm{1539} & \longrightarrow 2\times(\bm{1},\bm{1})_{0} \oplus \ldots\\
%   \end{aligned}
% \end{equation}
% where only representations featuring $\SO(5)\times\SO(3)$ singlets have been specified.

% \paragraph{\boldmath $\SO(4)\times\SO(4)$ singlets}
% Under $\E_{7(7)}\rightarrow\SL(4)\times\SL(4)\times\R_{+}$:
% \begin{equation}
%   \begin{aligned}
%     \bm{133} & \longrightarrow (\bm{1},\bm{1})_{0} \oplus (\bm{1},\bm{1})_{4} \oplus (\bm{1},\bm{1})_{-4} \oplus \ldots,\\
%     \bm{1539} & \longrightarrow 2\times(\bm{1},\bm{1})_{0} \oplus \ldots\\
%   \end{aligned}
% \end{equation}
% where only representations featuring $\SO(4)\times\SO(4)$ singlets have been specified.

% \paragraph*{} The singlets always come from the $\bm{63}$ and $\bm{720}$ of $\SL(8)_{\rm IIB}$ (and $\bm{70}$ for $\SL(4)\times\SL(4)\times\R_{+}$).

\section{\texorpdfstring{$S^{1}$}{S1} reductions of 4d dyonic gaugings}
Dyonic truncations from type II supergravities to four-dimensional gauged supergravity have been constructed in ref.~\cite{Inverso:2016eet} with $\E_{7(7)}$ exceptional field theory. Here, we further compactify on a circle and study whether it is possible to turn on additional gauging in three dimensions. Schematically:
% \begin{equation}
%   \text{Type IIA/B sugra} \xlongrightarrow[\SL(8-p)\times\SL(p)]{S^{7-p}\times S^{p-1}} \text{4d gauged sugra} \xlongrightarrow[\SL(2)]{S^{1}} \text{3d gauged sugra}.
% \end{equation}
\begin{equation}
  \begin{aligned}
    \begin{tikzpicture}
    \draw (0,0) node (10d) [align=center] {Type IIA/B \\ supergravity};
    \draw ($(10d)+(6,0)$) node (4d) [align=center] {4d gauged \\ supergravity};
    \draw ($(4d)+(4,0)$) node (3d) [align=center] {3d gauged \\ supergravity};

    \draw [thick, -stealth] (10d) -- (4d) node [midway, above] {\small $S^{7-p}\times S^{p-1}$} node [midway, below] {\small $\SL(8-p)\times \SL(p)$};
    \draw [thick, -stealth] (4d) -- (3d) node [midway, above] {\small $S^{1}$} node [midway, below] {\small $\SL(2)$};
    \draw [thick, -stealth] ($(10d)-(0,0.6)$) arc [start angle=-160,end angle=-20,x radius=5.3,y radius=1.25] node [midway, above] {\small $S^{7-p}\times S^{p-1}\times S^{1}$} node [midway, below] {\small $\SL(8-p)\times \SL(p)\times \SL(2)$};
    \end{tikzpicture}
  \end{aligned}
\end{equation}
The truncation from 10d to 3d is constructed using $\E_{8(8)}$ exceptional field theory, with isometries embedded as $\SL(8-p)\times\SL(p)\times\SL(2) \subset \E_{7(7)} \times \SL(2) \subset \E_{8(8)}$. Under $\E_{8(8)}\rightarrow\E_{7(7)}\times\SL(2)$, the ExFT coordinates decompose as
\begin{equation}
  \bm{248} \longrightarrow (\bm{1},\bm{3}) \oplus (\bm{56},\bm{2}) \oplus (\bm{133},\bm{3}).
\end{equation}
The $\E_{7(7)}$ coordinates sit in $(\bm{56},\bm{2})$. \ce{Further review ref.~\cite{Inverso:2016eet}? At least coordinates and cie. Details of IIA and IIB embeddings of forms?}

  \subsection{\texorpdfstring{$S^{3}\times S^{3}\times S^{1}$}{S3xS3xS1}}
  In ref.~\cite{Inverso:2016eet}, a truncation from type IIB to 4d on $S^{3}\times S^{3}\times S^{1}$ has been constructed. The type IIB two-forms $\hat{C}^{\alpha}_{\hat{\mu}\hat{\nu}}$ ($\alpha$ denotes a $\SL(2)_{\rm IIB}$ doublet) lead to seven-form field-strengths $H_{(7)}^{\alpha}=\d C_{(6)}^{\alpha}$. We search for singlets within $H_{(7)}^{\alpha}$ on $S^{3}\times S^{3}\times S^{1}$. Coordinates $\{y^{m}\}$ on $S^{3}\times S^{3}$, $z$ on $S^{1}$. Choice of gauge such that $H_{(7)}^{\alpha}$ given by
  \begin{equation}
    \partial_{z}C_{mnpqrs}^{\alpha}.
  \end{equation}
  Thus, we have to search for singlets in $C_{(6)}^{\alpha}$, built on $S^{3}\times S^{3}$ only. \ce{Is the choice of gauge always possible?}

  We first have to identify how $\SL(2)_{\rm IIB}$ is embedded in $\E_{7(7)}$ (which will help us to identify $\SL(2)_{\rm IIB}$ doublets like $C_{(6)}^{\alpha}$). To do so, we first follow ref.~\cite{Inverso:2016eet}, where the internal coordinates $\{y^{m}\}$ are embedded into $\E_{7(7)}$ following the decomposition $\E_{7(7)}\longrightarrow\SL(6)\times\SL(2)_{\rm IIB}\times\R^{+}_{\rm IIB}$:
  \begin{equation}
    \bm{56} \rightarrow (\bm{6'},\bm{1})_{-4} \oplus (\bm{6},\bm{2})_{-2} \oplus (\bm{20},\bm{1})_{0} \oplus (\bm{6'},\bm{2})_{2} \oplus (\bm{6},\bm{1})_{4},
  \end{equation}
  with coordinates in $(\bm{6'},\bm{1})_{-4}$. To build the truncation on $S^{3}\times S^{3}$, we are rather interested in the decomposition under $\E_{7(7)}\longrightarrow\SL(4)_{1}\times\SL(4)_{2}\times\R^{+}_{\SL(8)}$ (through $\SL(8)$):
  \begin{equation}
    \bm{56} \rightarrow \left[(\bm{6},\bm{1})_{2} \oplus (\bm{4},\bm{4})_{0} \oplus (\bm{1},\bm{6})_{-2}\right] \oplus \left[(\bm{1},\bm{6})_{2} \oplus (\bm{\bar{4}},\bm{\bar{4}})_{0} \oplus (\bm{6},\bm{1})_{-2}\right].
  \end{equation}
  We compare these two decompositions by further decomposing under
  \begin{equation}
    \begin{aligned}
     (i):&\qquad \E_{7(7)}\longrightarrow\SL(6)\times\SL(2)_{\rm IIB}\times\R^{+}_{\rm IIB} \longrightarrow \SL(3)_{1}\times\SL(3)_{2}\times\R^{+}_{\SL(6)}\times\R^{+}_{\SL(2)_{\rm IIB}}\times\R^{+}_{\rm IIB},\\
     (ii):&\qquad \E_{7(7)}\longrightarrow\SL(4)_{1}\times\SL(4)_{2}\times\R^{+}_{\SL(8)} \longrightarrow \SL(3)_{1}\times\SL(3)_{2}\times\R^{+}_{1}\times\R^{+}_{2}\times\R^{+}_{\SL(8)}.
    \end{aligned}
  \end{equation}
  We get the following:
  \begin{equation}
    \begin{aligned}
    \bm{56} &\overset{(i)}{\longrightarrow} \Big[(\bm{\bar{3}},\bm{1})_{-1,0,-4} \oplus (\bm{1},\bm{\bar{3}})_{1,0,-4}\Big] \oplus \Big[(\bm{3},\bm{1})_{1,1,-2} \oplus (\bm{3},\bm{1})_{1,-1,-2} \oplus (\bm{1},\bm{3})_{-1,1,-2} \oplus (\bm{1},\bm{3})_{-1,-1,-2}\Big] \\
    &\qquad \oplus \Big[(\bm{1},\bm{1})_{3,0,0} \oplus (\bm{1},\bm{1})_{-3,0,0} \oplus (\bm{3},\bm{\bar{3}})_{-1,0,0} \oplus (\bm{\bar{3}},\bm{3})_{1,0,0}\Big] \\
    &\qquad \oplus \Big[(\bm{\bar{3}},\bm{1})_{-1,1,2} \oplus (\bm{\bar{3}},\bm{1})_{-1,-1,2} \oplus (\bm{1},\bm{\bar{3}})_{1,1,2} \oplus (\bm{1},\bm{\bar{3}})_{1,-1,2}\Big] \oplus \Big[(\bm{3},\bm{1})_{1,0,4} \oplus (\bm{1},\bm{3})_{-1,0,4}\Big],\\[10pt]
    \bm{56} &\overset{(ii)}{\longrightarrow} \Big[(\bm{\bar{3}},\bm{1})_{2,0,2} \oplus (\bm{3},\bm{1})_{-2,0,2} \oplus (\bm{1},\bm{1})_{-3,-3,0} \oplus (\bm{1},\bm{3})_{-3,1,0}\oplus (\bm{3},\bm{1})_{1,-3,0} \oplus (\bm{3},\bm{3})_{1,1,0}\\
      &\quad\qquad \oplus (\bm{1},\bm{\bar{3}})_{0,2,-2} \oplus (\bm{1},\bm{3})_{0,-2,-2}\Big] \oplus \Big[(\bm{1},\bm{\bar{3}})_{0,2,2} \oplus (\bm{1},\bm{3})_{0,-2,2} \oplus (\bm{1},\bm{1})_{3,3,0} \oplus (\bm{1},\bm{\bar{3}})_{3,-1,0}\\
      &\quad\qquad \oplus (\bm{\bar{3}},\bm{1})_{-1,3,0} \oplus (\bm{\bar{3}},\bm{\bar{3}})_{-1,-1,0} \oplus (\bm{\bar{3}},\bm{1})_{2,0,-2} \oplus (\bm{3},\bm{1})_{-2,0,-2}\Big].
    \end{aligned}
  \end{equation}

  From these decompositions we extract the relations between the charges:
  \begin{equation}
    \begin{cases}
    q_{\SL(6)} = -\dfrac{1}{2}(q_{\SL(3)_{1}}+q_{\SL(3)_{2}}), \\[7pt]
    q_{\SL(2)_{\rm IIB}} = \dfrac{1}{4}\,\left(q_{\SL(3)_{1}}-q_{\SL(3)_{2}}+q_{\SL(8)}\right), \\[7pt]
    q_{\rm IIB} = \dfrac{1}{2}\,\left(-q_{\SL(3)_{1}}+q_{\SL(3)_{2}}+3\,q_{\SL(8)}\right).
    \end{cases}
  \end{equation}
  So the coordinates are inside $(\bm{6},\bm{1})_{-2}\oplus(\bm{1},\bm{6})_{2}$ of $\SL(4)_{1}\times\SL(4)_{2}\times\R^{+}_{\SL(8)}$. \ce{Problem with $3, \bar{3}$? Comes from identification $\bm{6'}=\bm{\bar{6}}$?}

  We now turn to the six-form potential $C_{(6)}^{\alpha}$, which sits in the representation $\bm{133}$ of $\E_{7(7)}$. Under $\E_{7(7)} \rightarrow \SL(8)$, it decomposes as $\bm{133} \rightarrow \bm{63} \oplus \bm{70}$, and
  \begin{equation}
    \begin{aligned}
    \SL(8) & \longrightarrow \SL(4) \times \SL(4) \times \R^{+} \\
    \bm{63} & \longrightarrow (\bm{4},\bm{\bar{4}})_{2} \oplus (\bm{1},\bm{1})_{0} \oplus (\bm{1},\bm{15})_{0} \oplus (\bm{15},\bm{1})_{0} \oplus (\bm{\bar{4}},\bm{4})_{-2} \\
    \bm{70} & \longrightarrow (\bm{1},\bm{1})_{4} \oplus (\bm{\bar{4}},\bm{4})_{2} \oplus (\bm{6},\bm{6})_{0} \oplus (\bm{4},\bm{\bar{4}})_{-2} \oplus (\bm{1},\bm{1})_{-4}.
    \end{aligned}
  \end{equation}
  Given that $C_{(6)}^{\alpha}$ is a $\SL(2)_{\rm IIB}$ doublet of fixed $R_{\rm IIB}^{+}$ charge, let us spell out the charges explicitly. In $\bm{63}$:
  \begin{equation}
    \begin{aligned}
    \SL(4) \times \SL(4) \times \R^{+} & \longrightarrow \SL(3) \times \SL(3) \times \R^{+}_{\SL(2)_{\rm IIB}} \times \R^{+}_{\rm IIB} \\
    (\bm{4},\bm{\bar{4}})_{2} & \longrightarrow (\bm{1},\bm{1})_{-1,6} \oplus (\bm{1},\bm{\bar{3}})_{0,4} \oplus (\bm{3},\bm{1})_{0,4} \oplus (\bm{3},\bm{\bar{3}})_{1,2}\\
    (\bm{1},\bm{1})_{0} & \longrightarrow (\bm{1},\bm{1})_{0,0} \\
    (\bm{1},\bm{15})_{0} & \longrightarrow (\bm{1},\bm{1})_{0,0} \oplus (\bm{1},\bm{3})_{-1,2} \oplus (\bm{1},\bm{\bar{3}})_{1,-2} \oplus (\bm{1},\bm{8})_{0,0}\\
    (\bm{15},\bm{1})_{0} & \longrightarrow (\bm{1},\bm{1})_{0,0} \oplus (\bm{3},\bm{1})_{1,-2} \oplus (\bm{\bar{3}},\bm{1})_{-1,2} \oplus (\bm{8},\bm{1})_{0,0} \\
    (\bm{\bar{4}},\bm{4})_{-2} & \longrightarrow (\bm{1},\bm{1})_{1,-6} \oplus (\bm{1},\bm{3})_{0,-4} \oplus (\bm{\bar{3}},\bm{1})_{0,-4} \oplus (\bm{\bar{3}},\bm{3})_{-1,-2}.
    \end{aligned}
  \end{equation}
  In $\bm{70}$:
  \begin{equation}
    \begin{aligned}
    \SL(4) \times \SL(4) \times \R^{+} & \longrightarrow \SL(3) \times \SL(3) \times \R^{+}_{\SL(2)_{\rm IIB}} \times \R^{+}_{\rm IIB} \\
    (\bm{1},\bm{1})_{4} & \longrightarrow (\bm{1},\bm{1})_{1,6}\\
    (\bm{\bar{4}},\bm{4})_{2} & \longrightarrow (\bm{1},\bm{1})_{2,0} \oplus (\bm{1},\bm{3})_{1,2} \oplus (\bm{\bar{3}},\bm{1})_{1,2} \oplus (\bm{\bar{3}},\bm{3})_{0,4} \\
    (\bm{6},\bm{6})_{0} & \longrightarrow (\bm{3},\bm{3})_{0,0} \oplus (\bm{3},\bm{\bar{3}})_{-1,2} \oplus (\bm{\bar{3}},\bm{3})_{1,-2} \oplus (\bm{\bar{3}},\bm{\bar{3}})_{0,0} \\
    (\bm{4},\bm{\bar{4}})_{-2} & \longrightarrow (\bm{1},\bm{1})_{-2,0} \oplus (\bm{1},\bm{\bar{3}})_{-1,-2} \oplus (\bm{3},\bm{1})_{-1,-2} \oplus (\bm{3},\bm{\bar{3}})_{0,-4} \\
    (\bm{1},\bm{1})_{-4} & \longrightarrow (\bm{1},\bm{1})_{-1,-6}
    \end{aligned}
  \end{equation}
  $\SO(3)\times\SO(3)$ singlets $\SL(2)_{\rm IIB}$ doublets and of given $R_{\rm IIB}^{+}$ charge are given by
  \begin{equation}
    (\bm{1},\bm{1})_{1,6} \oplus (\bm{1},\bm{1})_{-1,6} \subset (\bm{1},\bm{1})_{4} \oplus (\bm{4}, \bm{\bar{4}})_{2} \quad {\rm and} \quad (\bm{1},\bm{1})_{1,-6} \oplus (\bm{1},\bm{1})_{-1,-6} \subset (\bm{1},\bm{1})_{-4} \oplus (\bm{\bar{4}}, \bm{4})_{-2},
  \end{equation}
  where we wrote embeddings of $\SL(3) \times \SL(3) \times \R^{+}_{\SL(2)_{\rm IIB}} \times \R^{+}_{\rm IIB}$ in $\SL(4) \times \SL(4) \times \R^{+}$. These doublets correspond to $C_{(6)}^{\alpha}$ and the analogous $C^{(6)\,\alpha}$. So
  \begin{equation}
    C_{(6)}^{\alpha} \subset (\bm{1},\bm{1})_{4} \oplus (\bm{4},\bm{\bar{4}})_{2}.
  \end{equation}
  \ce{But no $(\bm{4},\bm{\bar{4}})$ in $\left((\bm{6},\bm{1})_{2}\oplus(\bm{1},\bm{6})_{-2}\right)^{\wedge 6}$...}

  \subsection{\texorpdfstring{$S^{7-p}\times S^{p-1}$}{S7-pxSp-1}}
  ...
